---
title: "Data Analysis"
author: "Nnenna Anako, Joanne Song, Nicandro Mandujano Acevedo"
date: "4/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing Data and loading packages
```{r}
DHSdata <- read.csv("NewADAProject.csv")

library(dplyr)
library(ggplot2)
library(car)
library(tidyverse)
```

### Renaming Variables
```{r}
# Rename variables of who perpetuates violence
DHSdataC<-rename(DHSdata, Age=v012, Age_Cat=v013, WealthIndex=v190, WealthScore=v191, 
                 UnionStatus=v502, EmoVio=d104, LessPA=d106, 
                 SeverPA=d107, SexVio=d108, HouseholdLiving=hv012, SonsLivingH=v202, 
                 DaughtersLivingH=v203,Working=v714, TypeRes=v140, PartDrink=d113)
```

### Recode a variable for total children living at home
```{r}
DHSdataC<-dplyr::mutate(DHSdataC, ChildHouseLive = SonsLivingH + DaughtersLivingH)
```

### Creating a new Variable for those that have experienced any form of violenve
```{r}
#Firdst need to make this a factor
DHSdataC$EmoVio <- as.factor(DHSdataC$EmoVio)
DHSdataC$LessPA <- as.factor(DHSdataC$LessPA)
DHSdataC$SeverPA <- as.factor(DHSdataC$SeverPA)
DHSdataC$SexVio <- as.factor(DHSdataC$SexVio)

# Make all 9 be NAs
DHSdataC$EmoVio<-dplyr::na_if(DHSdataC$EmoVio, 9)
DHSdataC$LessPA<-dplyr::na_if(DHSdataC$LessPA, 9)
DHSdataC$SexVio<-dplyr::na_if(DHSdataC$SexVio, 9)
DHSdataC$SeverPA<-dplyr::na_if(DHSdataC$SeverPA, 9)

# Make varible for if any type of violence
DHSdataC$AnyVio <- ifelse(DHSdataC$EmoVio==1|DHSdataC$LessPA==1|DHSdataC$SeverPA==1|
                            DHSdataC$SexVio==1,1,0)

# Change yo factor
DHSdataC$AnyVio<-as.factor(DHSdataC$AnyVio)

#Plotting to make sure it worked
DHSdataC %>% 
  filter(!is.na(AnyVio)) %>% 
ggplot(aes(x=AnyVio))+
  geom_bar(na.rm = T)
```

# RUNNING LOGISTIC REGRESSIONS
Four models to be run:
1- Children living in household plus controls
2- Work plus control
3- Partner drinks plus control
4- All independent variables controls

### Model 1 - Children living in the household as IV

1. Examining the distribution of children living in the household by ever expereinced violence
```{r}
#Make labels for ever expereinced violence
DHSdataC$AnyVio<- factor(DHSdataC$AnyVio, labels = c("Never Experienced", "Experienced Any"))

#Make box-plot 
DHSdataC %>%
  drop_na(c(AnyVio, ChildHouseLive)) %>%
  ggplot(aes(x = AnyVio, y = ChildHouseLive)) +
  geom_boxplot(aes(fill = AnyVio)) +
  labs(x = "Any Violence", y = "Children Living in Household") +
  theme_bw()
```

__Based on the resulting plot, it appears that the distribution of children at home does not vary by ever experienced violence__

2. Determining if running age and number children living in the household can be run as continuous variables

```{r}
#linearity for age
age.times.logage <- DHSdataC$Age * log(DHSdataC$Age)#create term to test linearity

boxTidwellAge <- glm(AnyVio ~ Age + age.times.logage, data=DHSdataC, family="binomial") #Box Tidwell technique, test the assumption of linearity

summary(boxTidwellAge)

#linearity for children
child.times.logchild <- DHSdataC$ChildHouseLive * log(DHSdataC$ChildHouseLive)#create term to test linearity

boxTidwellChild <- glm(AnyVio ~ ChildHouseLive + child.times.logchild, data=DHSdataC, family="binomial") #Box Tidwell technique, test the assumption of linearity

summary(boxTidwellChild)
```

__Based on the box Tidwell results, neither the age or children living at home variables should be run a continuous. Therefore categorical variables need to be created for children living at home. Age already has a categorical varaible__

A. Creating a categorical variable for children living in the home
```{r}
#Creating a categorical variable for children living at home
DHSdataC$ChildHouse_Cat<-car::recode(DHSdataC$ChildHouseLive, "0='0'; 1:2='1-2'; 3:4='3-4'; else='5 or more'")

# Checking if categorical variable worked
ggplot(DHSdataC, aes(x=ChildHouse_Cat))+ geom_bar ()

# Factor and relabel
DHSdataC$ChildHouse_Cat<-as.factor(DHSdataC$ChildHouse_Cat)

# Visualizing that the age categorical variable is correct
ggplot(DHSdataC, aes(x=Age_Cat))+ geom_bar ()

# CHeck to see if the are any missing variables
table(DHSdataC$ChildHouse_Cat)
table(DHSdataC$Age_Cat)
```

3. Running Univariate logictic regeression model with only the number of children living at home
```{r}
# Children living at home logistic model
ChildLogit <- glm(AnyVio ~ ChildHouse_Cat, data=DHSdataC, family="binomial")
summary(ChildLogit)
  
# Calculate and print ORs and 95% CIs  
ORChild<-exp(cbind(OR = coef(ChildLogit), confint(ChildLogit))) #calculate ORs and 95% CIs
ORChild #print ORs and 95% CIs

```

_None are significant_

4. Running a multavariate model
```{r}
# Checking to see what is the class of the other categorical variables
class(DHSdataC$WealthIndex)
class(DHSdataC$TypeRes)

# Since all variables are factors, model can go ahead and be run
# Multivariate model
ChildMLogit <- glm(AnyVio ~ ChildHouse_Cat + Age_Cat + WealthIndex + TypeRes, data=DHSdataC, family="binomial")
summary(ChildMLogit)
  
# Calculate and print ORs and 95% CIs  
ORChildM<-exp(cbind(OR = coef(ChildMLogit), confint(ChildMLogit))) #calculate ORs and 95% CIs
ORChildM #print ORs and 95% CIs

```

_No children cat are significant_

5. Looking at influential cases

```{r}
#Create a graph of CD of the points/observations in the data to help guide which points may have undue influence in the model
plot(ChildMLogit, which=4, id.n=5, col="red") 

```

NOT SURE HOW TO ACTUALLY SEE WHAT ID THAT IS!!!!!

## Model 2 - Whether women currently working

1. Examining the distribution of women currently working by ever expereinced violence
```{r}
# Since these are categorical variables a two way table would be best look at the distribution
DHSdataC$Working<-dplyr::na_if(DHSdataC$Working, 9)

DHSdataC$Working<-car::recode(DHSdataC$Working, "0='Not Working'; 1='Currently Working'")

table(DHSdataC$AnyVio, DHSdataC$Working)
```

2. Running Univariate logictic regeression model with only working status of women
```{r}
# Working status logistic model
WorkingLogit <- glm(AnyVio ~ Working, data=DHSdataC, family="binomial")
summary(WorkingLogit)
  
# Calculate and print ORs and 95% CIs  
ORWorking<-exp(cbind(OR = coef(WorkingLogit), confint(WorkingLogit))) #calculate ORs and 95% CIs
ORWorking #print ORs and 95% CIs

```

_Working status not significant_

3. Running a multavariate model for working status
```{r}
# Multivariate model
WorkingMLogit <- glm(AnyVio ~ Working + Age_Cat + WealthIndex + TypeRes, data=DHSdataC, family="binomial")
summary(WorkingMLogit)
  
# Calculate and print ORs and 95% CIs  
ORWorkingM<-exp(cbind(OR = coef(WorkingMLogit), confint(WorkingMLogit))) #calculate ORs and 95% CIs
ORWorkingM #print ORs and 95% CIs
```
_Not working become protective against violence_

4. Looking at influential cases

```{r}
#Create a graph of CD of the points/observations in the data to help guide which points may have undue influence in the model
plot(WorkingMLogit, which=4, id.n=5, col="red") 

```

NOT SURE HOW TO FIGURE OUT WHICH ID THAT IS EITHER!!!

## Model 3 - Partner drinking

1. Examining the distribution of partner drinking status by ever expereinced violence
```{r}
# Since these are categorical variables a two way table would be best look at the distribution
class(DHSdataC$PartDrink)
DHSdataC$PartDrink<=as.factor(DHSdataC$PartDrink)

DHSdataC$PartDrink<-dplyr::na_if(DHSdataC$PartDrink, 8)
DHSdataC$PartDrink<-dplyr::na_if(DHSdataC$PartDrink, 9)

DHSdataC$PartDrink<-car::recode(DHSdataC$PartDrink, "0='Not Drinker'; 1='Drinker'")

table(DHSdataC$AnyVio, DHSdataC$PartDrink)

# Setting "Not drinker" to be level 0
DHSdataC$PartDrink[DHSdataC$PartDrink=='Not Drinker']<-0
```
_Based on this there might be a ralation between whether partner drinks and experiencing any violence

2. Running Univariate logictic regeression model with partner drinking status
```{r}
# Partner Drink status logistic model
DrinkLogit <- glm(AnyVio ~ PartDrink, data=DHSdataC, family="binomial")
summary(DrinkLogit)
  
# Calculate and print ORs and 95% CIs  
ORDrink<-exp(cbind(OR = coef(DrinkLogit), confint(DrinkLogit))) #calculate ORs and 95% CIs
ORDrink #print ORs and 95% CIs

```
_Partner not being a drinker is a protective factor_

3. Running a multavariate model for partner drinking status
```{r}
# Multivariate model
DrinkMLogit <- glm(AnyVio ~ PartDrink + Age_Cat + WealthIndex + TypeRes, data=DHSdataC, family="binomial")
summary(DrinkMLogit)
  
# Calculate and print ORs and 95% CIs  
ORDrinkM<-exp(cbind(OR = coef(DrinkMLogit), confint(DrinkMLogit))) #calculate ORs and 95% CIs
ORDrinkM #print ORs and 95% CIs
```

_Protective effect slightly increases_

4. Looking at influential cases

```{r}
#Create a graph of CD of the points/observations in the data to help guide which points may have undue influence in the model
plot(DrinkMLogit, which=4, id.n=5, col="red") 

```

AGAIN NOT SURE HOW TO GET THE ID OF THAT SPECIFIC ID

## 4 - Running model with all IV variables

1. Running model with only the three IV of interests

```{r}
# Only IV
AllLogit <- glm(AnyVio ~ ChildHouse_Cat + Working + PartDrink, data=DHSdataC, family="binomial")
summary(AllLogit)
  
# Calculate and print ORs and 95% CIs  
ORAll<-exp(cbind(OR = coef(AllLogit), confint(AllLogit))) #calculate ORs and 95% CIs
ORAll #print ORs and 95% CIs
```

_Two significant predictor is partner not drinking (protective), and having more than 5 children (risk)_

2. Running final model with all IV and controls
```{r}
# IV and controls
FinalLogit <- glm(AnyVio ~ ChildHouse_Cat + Working + PartDrink + Age_Cat + WealthIndex + TypeRes, data=DHSdataC, family="binomial")
summary(FinalLogit)
  
# Calculate and print ORs and 95% CIs  
ORFinal<-exp(cbind(OR = coef(FinalLogit), confint(FinalLogit))) #calculate ORs and 95% CIs
ORFinal #print ORs and 95% CIs
```

_Not working becomes protective, more than 5 children risk, not drinker protective_

3. Looking at influential cases

```{r}
#Create a graph of CD of the points/observations in the data to help guide which points may have undue influence in the model
plot(FinalLogit, which=4, id.n=5, col="red") 

```

AGAIN NOT SURE HOW TO GET THE ID FOR THIS CASE?


**********************************************
Code for getting metrics for Table 1
**********************************************
```{r}
# First need full analytical sample
DHSdataFull<- filter(DHSdataC, !is.na(AnyVio))
DHSdataFull2<- filter(DHSdataC, !is.na(AnyVio) & !is.na(Age_Cat) & !is.na(WealthIndex) & !is.na(TypeRes) & !is.na(ChildHouse_Cat) & !is.na(Working) & !is.na(PartDrink))
```
```{r}
# Getting values
table(DHSdataFull2$Age_Cat) # Age
table(DHSdataFull2$AnyVio) # Violence
table(DHSdataFull2$WealthIndex) # Wealth Index
table(DHSdataFull2$ChildHouse_Cat) # Number of children living in the household
table(DHSdataFull2$Working) # Employment status
table(DHSdataFull2$PartDrink) # Spouse drinking status
table(DHSdataFull2$TypeRes) # Type of residence

```
